---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Smoothing tracks and aggregating dat

## Prepare libraries {-}

```{r prep_libs_02_01}
library(data.table)
library(atlastools)
library(ggplot2)
```


## Reading in data {-}

Here we read the simulated data in. The limits of the coordinates are 0 and 1, so we add a few extraneous coordinates within a plausible distance to demonstrate the basic forms of filtering. We also assign a fictional _NBS_ value ranging between 2 and 6 to demonstrate filtering on covariates. Finally, we introduce a 'reflection', a phenomenon that results when a TOA system's position likelihood estimator shifts the real position by a large distance for an extended duration.

```{r read_sim_data_2}
# read in the data
data <- fread("data/data_sim.csv")[5000:10000, ]
data[, window_size := NA]
# data with small scale errors but no reflections or outliers
data_errors <- fread("data/data_no_reflection.csv")
data_errors[, window_size := 0]
```

## Applying a median smooth

```{r}
# smooth the data
list_of_smooths <- lapply(c(3, 5, 11, 21), function(z) {
  data_copy <- copy(data_errors)
  data_copy <- atl_median_smooth(data = data_copy,
                    x = "x", 
                    y = "y",
                    time = "time",
                    moving_window = z)
  data_copy[, window_size := z]
})
```

```{r}
# prepare data to plot
# make list of data to plot
data_plot <- append(list(data, data_errors), list_of_smooths)

# add offset
data_plot <- mapply(FUN = function(df, offset) {
  df[, x := x + offset]
}, data_plot, seq(from = 0, to = 1, length.out = length(data_plot)),
SIMPLIFY = FALSE)

# bind rows
data_plot <- rbindlist(data_plot, fill = TRUE)

plot_figure <- ggplot()+
  # geom_point(data = data_plot,
  #            aes(x, y),
  #            col = "darkgrey", 
  #            alpha = 1,
  #            shape = 1)+
  geom_path(data = data_plot,
            aes(x, y, 
                col = window_size,
                group = window_size),
            alpha = 1,
            show.legend = FALSE)+
  annotate(x = 0.6 + seq(0, 1, length.out = 6),
           y = 1.05,
           label = glue::glue('({letters[seq(6)]})'),
           geom = "text",
           fontface = "bold")+
  scale_colour_viridis_c(option = "D",
                         na.value = "red")+
  coord_equal() +
  theme_test()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())

plot_figure

# save figure
ggsave(plot_figure, filename = "figures/fig_median_smooth.png",
       width = 185, height = 185 * 2 / 3, units = "mm")
```

## Aggregating data

```{r}
# choose the data
data_agg <- copy(data)

list_of_agg <- lapply(c(10, 30, 120), function(z) {
  data_agg <- atl_thin_data(data = data_agg,
                            interval = z,
                            method = "aggregate")
  data_agg[, interval := z]
  
})
```

```{r}
# plot
data_plot <- append(list(data[, `:=`(SD = 0, interval = NA)]), list_of_agg)

# add offset
data_plot <- mapply(FUN = function(df, offset) {
  df[, x := x + offset]
}, data_plot, seq(from = 0, to = 1, length.out = length(data_plot)),
SIMPLIFY = FALSE)

data_plot <- rbindlist(data_plot, fill = TRUE)

plot_figure <- ggplot(data = data_plot)+
  geom_point(aes(x, y, size = SD),
             colour = "darkblue",
             alpha = 0.5,
             shape = 16, show.legend = F)+
  geom_path(aes(x, y,
                group = interval,
                col = interval),
            col = "red",
            show.legend = FALSE)+
  annotate(x = 0.6 + seq(0, 1, length.out = 4),
           y = 1.05,
           label = glue::glue('({letters[seq(4)]})'),
           geom = "text",
           fontface = "bold")+
  coord_equal()+
  theme_test()+
  theme(axis.text = element_blank(),
        axis.title = element_blank())

plot_figure

# save figure
ggsave(plot_figure, filename = "figures/fig_aggregate_data.png",
       width = 185, height = 90, units = "mm")
```

